---
title: "<span style='font-size: 40px'>The Association between Frailty and the Predictive Performance of ANZROD Score</style>"
author: "Ryo Ueno, Ashwin Subramaniam, David Pilcher" 
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output: 
        html_document:
                toc: true
---

```{r setup, include=FALSE}
require(tidyverse)
require(knitr)
require(kableExtra)
require(tableone)
require(pROC)
require(ModelMetrics)
require(ResourceSelection)
require(gbm)
rm(list=ls())
dt <- readRDS("./data/franzrod.obj")
```

For further details, please refer to this [online repository](https://github.com/ryo313/franzrod) or [contact author](mailto: ryo.ueno@monash.edu)

### Objectives of this report

- To clarify the motivation of the study
- To clarify the initial statistic plan
- To share the exploratory analysis results 

### Summary of the key findings 
- Both discrimination and calibration of ANZROD score differs amongst different frailty population 
- This result may support the future research to incorporate frailty score into ANZROD.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
read.csv("./tables_figures/table2.csv") %>%
        select(-1) %>% knitr::kable(caption = "Summary Table: Comparison of predictive values of ANZROD to predict in-hospital mortality in three cohorts") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = F)
```


### Background
- **Australian and New Zealand Risk of Death (ANZROD)** specific for Australian and New Zealand intensive care units (ICUs) has better calibration and discrimination in predicting mortality compared with the existing prognostic models such as APACHE III-j model.

- **Frailty** is a "physiological age" rather than a "chronological age". 

- Frailty is defined as a multidimensional syndrome characterised by a decline of physical, physiologic and cognitive reserves. Frailty increases with age and is characterised by a heterogeneous combination of decreased mobility, weakness, reduced muscle mass, poor nutritional status and diminished cognitive function. 

- **ANZROD does not factor in frailty**, which is associated with worse prognosis in critically ill patients. Frailty is an independent risk factor after adjustment for classic risks such as severity, age, sex, comorbidities. 

- We hypothesized that **ANZROD may underestimate the risk of mortality amongst frail populations**. 

### Methods

#### Study design
- **Study design:** A multicentre retrospective cohort study
- **Setting:** 147 ICUs in Australia and New Zealand
- **Inclusion:** All ICU admissions from 2017 to 2018 in the ANZICS-APD
- **Exclusion:** Patietns without Clinical Frailty Scale data
- **prediction outcome:** in-hospital mortality
- **Frailty:** frailty was measured using [Clinical Frailty Scale (CFS)](https://github.com/ryo313/franzrod/blob/master/others/ClinicalFrailtyScale.png)  

#### Statistic analysis
- All the patients were categorized in the following three groups
<br/>
        1. **non-frail (CFS = 1-3)**
<br/>
        2. **pre-frail (CFS = 4)**
<br/>
        3. **frail (CFS = 5-9)**

- In each cohort, we assessed the following predictive value of ANZROD to predict in-hospital mortality
<br/>
        1. **Area under the receiver operating characteristic curve (AUROC)**
<br/>
        2. **Brier score**
<br/>
        3. **Standardized Mortality Ratio (SMR)** 
<br/>
        4. **Hosmer-Lemeshow goodness-of-fit**

- We created the receiver operating characteristic curve and calibration curve to.
- please refer to the [supplementary document](https://github.com/ryo313/franzrod/blob/master/citations/Paul2013_supplementary/1-s2.0-S0883944113002566-mmc3.doc) for the details of each values. 

### Preliminary Analysis Results 

#### Table 1-1: Patient characteristics: continuous variables
- Frail patients were older and have higher ANZROD score and APACHE-IIIj score. 

```{r table1, include=FALSE}
dt %>% 
        select(age, sex,elect,medsurg, chr_resp:cirrhos, frailty,frail_CFS,
              anzrodriskofdeath,apache3score,apache3riskofdeath,died_icu,
              died_hosp,admsourc) -> dt_table 

dt_table %>% 
        tableone::CreateTableOne(vars = names(.), 
                                 data = ., 
                                 strata = "frail_CFS") -> tableone

## make a continuous variable table
contVars = c("age","anzrodriskofdeath","apache3score","apache3riskofdeath")
CreateContTable(vars = contVars, data = dt_table, strata = "frail_CFS") -> contTable
nonNormalVars <- c("anzrodriskofdeath","apache3score","apache3riskofdeath")
print(contTable, nonnormal = nonNormalVars, test = F) %>%
        write.csv(.,"./tables_figures/table1_cont.csv")

## make a categorical value table
tableone$CatTable %>% 
        print(.,test=F) %>% 
        write.csv(.,"./tables_figures/table1_cat.csv")
```


```{r echo=FALSE}
read.csv("./tables_figures/table1_cont_edit.csv") %>% 
        as.data.frame() %>% 
        select(-5,-6) %>% 
        knitr::kable(caption = "Table 1-1: patient characteristics")%>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = F)
```

#### Table 1-2: Patient characteristics: categorical variables
- Frail cohort has more female patients.
- Frail cohort has more emergency ICU admissions. 
- Frail cohort has more comoridities, especially chronic respiratory conditions. 
- Frail cohort had higher mortality both during ICU and hospital stay. 

```{r echo=FALSE}
read.csv("./tables_figures/table1_cat_edit.csv") %>% 
        as.data.frame() %>% 
        select(-5,-6) %>% 
        knitr::kable(caption = "Table 1-2: Patient characteristics")%>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = F)%>%
        add_indent(c(4:6,8:18))
```


####Table 2: Comparison of predictive values of ANZROD to predict in-hospital mortality in three cohorts
- ROC was lower amongst frail cohort compared with non-frail cohort
- Calibration values were similar amongst three categories. 
- needs proofreading for the code
```{r message=FALSE, warning=FALSE, include=FALSE}
roc_calculate <- function(data, i=3){
        data <- data %>% 
                filter(!is.na(died_hosp) & !is.na(anzrodriskofdeath)) %>%
                mutate(died_hosp_num = as.numeric(died_hosp)-1)
        ci<- pROC::roc(data$died_hosp, predictor = data$anzrodriskofdeath) %>% ci()
        brier <- ModelMetrics::brier(data$died_hosp_num, data$anzrodriskofdeath)
        HL_list <- ResourceSelection::hoslem.test(data$died_hosp_num, data$anzrodriskofdeath,20)
        data.frame(
                ROC = paste(median(ci) %>% round(i)),
                ROC_95CI = paste("[",quantile(ci, 0.05) %>% round(i),";",
                                        quantile(ci, 0.95) %>% round(i) ,"]"),
                Brier_score = brier %>% round(i),
                SMR = (sum(data$died_hosp_num)/sum(data$anzrodriskofdeath)) %>% round(i),
                Hosmer_Lemeshow = HL_list$statistic %>% as.numeric() %>% round(i)
        )
        }

dt %>% 
        group_by(frail_CFS) %>% 
        do(roc_calculate(.)) %>%
        print() -> tabletwo
tabletwo %>% write.csv(.,"./tables_figures/table2.csv")

```


```{r echo=FALSE}
tabletwo %>% 
        knitr::kable(caption = "Table 2: Comparison of predictive values of ANZROD to predict in-hospital mortality in three cohorts")%>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = F)
```

#### Figure 1: Calibration Curve of ANZROD to predict inhospital mortality in three cohorts
- Diagonal line represents the line of perfect calibration.
- Graphically, calibration performance is worse amongst frail population. 
- needs proofreading for the code

```{r figs, echo=FALSE, fig.cap="\\label{fig:figs}Figure 1:Calibration Curve of ANZROD to predict inhospital mortality in three cohorts"}
data <- dt %>% filter(frail_CFS == "non-frail")
data_pre <- dt %>% filter(frail_CFS == "pre-frail")
data_frail <- dt %>% filter(frail_CFS == "frail")
calib_plot <- function(data, replace = T, linecol = "black", df = 6, lty = 1){
        data %>% 
                filter(!is.na(died_hosp) & !is.na(anzrodriskofdeath)) %>%
                mutate(died_hosp_num = as.numeric(died_hosp)-1) -> data
        gbm::calibrate.plot(data$died_hosp, data$anzrodriskofdeath, 
                            distribution="bernoulli", replace = replace,
                       line.par = list(col = linecol), shade.col = "white",
                       shade.density = 1, rug.par = list(side = 1),
                       xlab = "ANZROD Predicted Risk of Inhospital Mortality", 
                       ylab = "Observed Inhospital Mortality", xlim = NULL,
                       ylim = NULL, knots = NULL, df = df, lty = lty)
}
calib_plot(data, df = 2)
calib_plot(data_pre, replace=F,linecol="darkgoldenrod", df = 2, lty = 2)
calib_plot(data_frail, replace=F,linecol="darkslategray4", df = 2, lty = 3)
legend("bottomright", 
       legend = c("non-frail", "pre-frail","frail"), 
       col = c("black","darkgoldenrod", "darkslategray4"),
       bty = "n", 
       cex = 1.1, 
       lty=1,
       title = "Line Types",
       text.col = c("black","darkgoldenrod", "darkslategray4"))
```


#### Figure 2: ROC of ANZROD to predict in-hospital mortality in three cohorts
- consistent with table 2, ROC shown that ANZROD performs worse in frail population compared with non-frial population. 
- AUROC of ANROD to predict in-hospital mortality were as follows: 
<br/>
        1. **non-frail:** 0.928 [ 0.916 ; 0.94 ]
<br/>
        2. **pre-frail:** 0.865 [ 0.841 ; 0.888 ]
<br/>
        3. **frail:** 0.841 [ 0.821 ; 0.861 ]
<br/>

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="\\label{fig:figs}Figure 2: ROC of ANZROD to predict in-hospital mortality in three cohorts"}
roc_plot <- function(data, add = F, lty=1, linecol = "black"){
        pROC::roc(data$died_hosp, predictor = data$anzrodriskofdeath) %>% 
                plot(.,add = add, lty=lty, col = linecol)
}
roc_plot(data)
roc_plot(data_pre, add = T, lty = 2, linecol = "darkgoldenrod")
roc_plot(data_frail, add = T, lty = 3, linecol ="darkslategray4")
legend("bottomright", 
       legend = c("non-frail", "pre-frail","frail"), 
       col = c("black","darkgoldenrod", "darkslategray4"),
       bty = "n", 
       cex = 1.1, 
       lty=c(1,2,3),
       title = "Line Types",
       text.col = c("black","darkgoldenrod", "darkslategray4"))
```


### Limitations
- selection bias (patients without CFS were excluded)
- this study includes only 5% of the original cohort used for the development of ANZROD
- the clinical significance of the observed difference is unknown
- This result may just reflect the non-collapsibility of AUC; however, both clinical and epidemiological evidence supports the independent association of frailty with mortality

<br/>
<br/>
- For further details, please refer to this [online repository](https://github.com/ryo313/franzrod) or [contact author](mailto: ryo.ueno@monash.edu)
